Emitter Source Geolocation from Imparted Rotor Blade Modulation

### 2016-11-17 Background.tex
edits for figures and equations finished, working on scope and noise

\section{Thesis Scope}
\textcolor{red}{ADD WHAT THIS THESIS IS AND POINT OUT WHAT IT IS NOT!}

@@

The resulting Doppler equation is described by

@@

The equation for modeling the scenario for a transmitter located at 90\textdegree \space azimuth is

@@

where $MaxDopplerProfile$ is an array of the maximum upper envelope values for each elevation sample, $MinDopplerProfile$ is an array of minimum lower envelope values for each elevation sample, and $\theta_{Limited}$ is the limited elevation angle range

@@

%formulate the geometric equations
The equation for modeling the scenario for a transmitter located at 270\textdegree \space azimuth is similar to the equations for calculating the 90\textdegree \space scenario where $f_{DopplerMax}$ is now calculated with a smaller reflection radius of 2m and $\theta_{Limited}$ is now calculated by

@@

The equation for modeling the scenario for a transmitter located at 180\textdegree \space azimuth is

@@

The equation for modeling the scenario for a transmitter located at 0\textdegree \space azimuth is

>>>>

\section{Thesis Scope}
\textcolor{red}{EDITED}

The scope of this thesis is to investigate whether it is possible to use the effect of RBM to geolocate a transmitter in ideal conditions. Under ideal conditions the received signal is deterministic based upon the propagation model and the initial conditions. This thesis does not intend to include every physical effect but is limited to the effects defined in the propagation model and by the assumptions made about the physical scene. The ray tracing software developed for the thesis is created for the sole purpose of understanding RBM. In order to determine the signals interaction with the moving rotor blade in an ideal environment, the ray traced scene is limited to the rotor blades, transmitter and receiver. The signal produced by the simulation is used to derive an algorithm that determines both the azimuth and elevation angle of a transmitter. The derived algorithm is then evaluated on both the deterministic signal to determine its azimuth and elevation accuracy. The algorithm is then tested again on signals in the presence of differing levels of noise to determine the algorithm's performance in a more realistic environment.

@@

The resulting Doppler equation is:

@@

The equation for modeling the scenario for a transmitter located at 90\textdegree \space azimuth is:

@@

where $MaxDopplerProfile$ is an array of the maximum upper profile values for each elevation sample, $MinDopplerProfile$ is an array of minimum lower profile values for each elevation sample, and $\theta_{Limited}$ is the limited elevation angle range:

@@

%formulate the geometric equations
The equation for modeling the scenario for a transmitter located at 270\textdegree \space azimuth is similar to the equations for calculating the 90\textdegree \space scenario where $f_{DopplerMax}$ is now calculated with a smaller reflection radius of 2m and $\theta_{Limited}$ is:

@@

The equation for modeling the scenario for a transmitter located at 180\textdegree \space azimuth is:

@@

The equation for modeling the scenario for a transmitter located at 0\textdegree \space azimuth is:

### 2016-11-17 Radio_Propagation_Simulation.tex
edits for figures and equations finished, working on scope and noise

\subsection{Free-Space Path Loss}
%describe path loss
Free-Space Path Loss is calculated using the Friis transmission and is incorporated into the model by keeping track of the distance a ray travels and adjusting its power by

@@

\subsubsection{Specular Reflection}
In the case of a perfect reflection the angle of the reflection is calculated by

@@

\subsubsection{Reflection Loss}
Power loss due to surface reflection is described through the following coefficient

@@

%formula description
where $\theta_i$ is the angle of incidence with the object's surface normal, $\varepsilon$ is the dielectric constant for that surface, and $\theta_t$ is the angle of refraction as described by Snell's law

@@

The final reflected power is then

@@

%multipath fading
\subsection{Fading}
The transmitted signal power experiences fading described by

\textcolor{red}{FIGURE OUT WHAT TO PUT HERE FOR FADING COEFFICIENT?}

The power can then be calculated using

\textcolor{red}{FADING POWER?}

The superposition of each ray's faded power becomes the received signal.

\subsection{Doppler Effect}
%define doppler effect
It is assumed that both the receiver and the transmitter are not moving therefore, the only Doppler imparted on the signal will be from the rotation of the blades. The associated Doppler is calculated using

@@

where $\lambda$ is the wavelength of the transmitted signal based off of the speed of light $c$ and its initial frequency $f_0$. The velocity in the direction of the transmitter is given by

@@

%formula description
where $r$ is the radius along the length of the blade where the reflection occurs. $\omega_r$ is the angular velocity of the rotor. $\vec{d_p}$ is the normalized vector that is perpendicular to the rotor in its direction of travel and $\vec{d_t}$ is the normalized vector that is in the direction of the transmitted ray. The velocity in the direction of the reflection is given by

@@

%figure of 2D airfoil
\begin{figure}
	\begin{center}
		\includegraphics[width=15cm]{images/radio_propagation/2d_airfoil.png}
		\caption{Airfoil Cross-Section}
		\label{fig:airfoil}
	\end{center}
\end{figure}

@@

%modular parameters
All aspects of the rotor blade are parameterized. Starting with the airfoil shape, which is defined by a set of 2D coordinates. The length and the number of rib sections can be configured along with the number of blades and corresponding altitude, as well as the RPM. This allows the scene to replicate any type of rotor configuration and position it at a specific altitude.

@@

\section{Ray Tracing}
%describe the core tracing engine and describe the flow of the program
The ray tracing simulation is conducted in the Monte Carlo fashion acting on the scene and propagation models. This method of raytracing is established for creating realistic images in \cite{Suffern2007}, \cite{Pharr2010}. The simulation is performed over successive iterations. Each iteration is a discrete element called a frame and each frame will be updated to represent the change in the position of the rotor blade.

\textcolor{red}{ADD MORE ABOUT RAY TRACING}

\subsection{Computation}
%flow of the program
The ray tracing engine first sets up the environment, creating the surfaces and placing the objects in the scene, then the number of frames that will be computed. The number of frames is determined by the number of rotations that the rotor makes within the simulation, the speed of the rotor in RPM and the sampling frequency of the receiver defined before the beginning of the simulation. This is described by

%frame rates equation
\begin{equation}
	Frames = \frac{60  f_s}{RPM} Rev
	\label{eqn:frames}
\end{equation}

where $Frames$ is the number of frames that will be computed in the simulation, $f_s$ is the sampling frequency of the receiver, $RPM$ is the revolutions per minuet of the rotor, and $Rev$ is the number of revolutions that the rotor makes in the simulation.

In each frame a set number of rays are casted into the scene in a uniformly distributed circle in the direction of the rotor and receiver using

@@

where $r_u$ and $\theta_u$ are uniformly distributed random variables from $0 \leq r_u \leq BladeLength^{2}$ and $ 0 \leq \theta_u \leq 2\pi$ respectively.

>>>>

\subsection{Free-Space Path Loss}
%describe path loss
Free-Space Path Loss is calculated using the Friis transmission and is incorporated into the model by keeping track of the distance a ray travels and adjusting its power:

@@

\subsubsection{Specular Reflection}
In the case of a perfect reflection the angle of the reflection is:

@@

\subsubsection{Reflection Loss}
Power loss due to surface reflection is described through the following coefficient:

@@

%formula description
where $\theta_i$ is the angle of incidence with the object's surface normal, $\varepsilon$ is the dielectric constant for that surface, and $\theta_t$ is the angle of refraction as described by Snell's law:

@@

The final reflected power is then:

@@

%multipath fading
\subsection{Fading}
\textcolor{red}{EDITED}

The transmitted signal power experiences fading due to multiple signal paths in the environment. The resulting signal that the receiver sees is a superposition of the multiple copies of the transmitted signal. The difference in signal's attenuation, phase, and delay cause destructive and constructive interference at the receiver. In the ray tracing simulation the received signal is calculated by keeping track of the distance each ray travels as well as the reflections, then finding the superposition of those signals which will produce the constructive and destructive interference.

\subsection{Doppler Effect}
%define doppler effect
It is assumed that both the receiver and the transmitter are not moving therefore, the only Doppler imparted on the signal will be from the rotation of the blades. The associated Doppler is:

@@

where $\lambda$ is the wavelength of the transmitted signal based off of the speed of light $c$ and its initial frequency $f_0$. The velocity in the direction of the transmitter is:

@@

%formula description
where $r$ is the radius along the length of the blade where the reflection occurs. $\omega_r$ is the angular velocity of the rotor. $\vec{d_p}$ is the normalized vector that is perpendicular to the rotor in its direction of travel and $\vec{d_t}$ is the normalized vector that is in the direction of the transmitted ray. The velocity in the direction of the reflection is:

@@

%figure of 2D airfoil
\begin{figure}
	\begin{center}
		\includegraphics[width=14cm]{images/radio_propagation/2d_airfoil.png}
		\caption{Airfoil Cross-Section}
		\label{fig:airfoil}
	\end{center}
\end{figure}

@@

%modular parameters
All aspects of the rotor blade are parameterized. Starting with the airfoil shape, which is defined by a set of 2D coordinates. The length and the number of rib sections can be configured along with the number of blades and corresponding altitude, as well as the $\omega_r$. This allows the scene to replicate any type of rotor configuration and position it at a specific altitude.

@@

\section{Ray Tracing}
\textcolor{red}{EDITED}

%describe the core tracing engine and describe the flow of the program
The ray tracing simulation is conducted in the Monte Carlo fashion acting on the previously described scene and propagation models. Ray tracing allows for the calculation of a particle or wave's path within an environment in order to produce photo realistic images. This method is established for creating images in \cite{Suffern2007}, \cite{Pharr2010} by tracing light throughout a virtual 3D environment of objects. Objects have physical characteristics that determine its shape along with how light rays interact with that object. The rays are then casted, in a distributed manner according to the light source, into the scene. The image is then built up by evaluating the color and luminance of the ray at each pixel. In the case of this simulation light is replaced with RF waves and are evaluated based on the propagation model described, and lights and cameras are replaced with transmitters and receivers respectively. The simulation is then performed over successive iterations as the objects move within the scene. Each iteration is a discrete element called a frame where each frame is updated to represent the change in the position of the rotor blade.

\subsection{Computation}
%flow of the program
The ray tracing engine first sets up the environment, creating the surfaces and placing the objects in the scene, then the number of frames that will be computed. The number of frames is determined by the number of rotations that the rotor makes within the simulation, the speed of the rotor $\omega_r$ and the sampling frequency of the receiver defined before the beginning of the simulation. The number of frames in the simulation is:

%frame rates equation
\begin{equation}
	n_{Frames} = \frac{f_s}{\omega_r}\omega_{Rev}
	\label{eqn:frames}
\end{equation}

where $n_{Frames}$ is the number of frames that will be computed in the simulation, $f_s$ is the sampling frequency of the receiver, $\omega_r$ is the rotational velocity of the rotor, and $\omega_{Rev}$ is the rotation that the rotor makes in the simulation.

In each frame a set number of rays are casted into the scene in a uniformly distributed circle in the direction of the rotor and receiver:

@@

where $r_u$ and $\theta_u$ are uniformly distributed random variables from $0 \leq r_u \leq r_b^{2}$ and $ 0 \leq \theta_u \leq 2\pi$ respectively.

### 2016-11-17 Results_new.tex
edits for figures and equations finished, working on scope and noise

\section{Introduction}
%introduce the results section and what processing will be done
The information gathered from the previous section is used to estimate the azimuth and elevation angles of the transmitter. First the azimuth angle will be estimated which will provide the information necessary for also calculating the elevation angle. The results of the estimation techniques are then analyzed to determine their accuracy.

@@

\section{Transmitter Azimuth Angle Estimation} \label{sec:transmitter_azimuth_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
The azimuth angle estimation technique is based on the data collected by rotating the helicopter platform in a clockwise manner, during which a Doppler envelope and subsequent maximum and minimum values are calculated. After a complete rotation the Doppler maximum and minimum values produce a sinusoidal pattern as shown in Figure \ref{fig:tx_azimuth_angle_200} without knowing the starting azimuth angle. With this data we find an unambiguous angle that will set the coordinate space and provide an estimate of current azimuth angle. To do this the absolute value of the minimum envelope data is superimposed onto the maximum envelope data as shown in Figure \ref{fig:azimuth_estimation_max_vs_absMin}.

\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/results/Azimuth_angle_estimation_max_vs_absMin.eps}
		\caption{Max and Absolute value of Min Envelope Frequencies vs Transmitter Azimuth Angle at an Elevation Angle of 45\textdegree}
		\label{fig:azimuth_estimation_max_vs_absMin}
	\end{center}
\end{figure}

From Figure \ref{fig:azimuth_estimation_max_vs_absMin} there are two points at which the maximums and the absolute value of the minimums cross. We then take the difference

%formula for difference
\begin{equation}
	 Envelope Difference = Max Upper Envelope - | Min Lower Envelope |
	 \label{eqn:difference}
\end{equation}

where $Max Upper Envelope$ is an array of the maximum value of each upper envelope calculation, and $Min Lower Envelope$ is an array of the minimum value of each lower envelope calculation. The result is an almost exact sinusoid centered around zero, shown in Figure \ref{fig:azimuth_estimation_difference}.

 \begin{figure}[b]
	\begin{center}
		\includegraphics[width=10cm]{images/results/Azimuth_angle_estimation_difference.eps}
		\caption{Difference between Max and Min Envelope Calculations vs Transmitter Azimuth Angle at an Elevation Angle of 45\textdegree}
		\label{fig:azimuth_estimation_difference}
	\end{center}
\end{figure}

While Figure \ref{fig:azimuth_estimation_difference} provides information about the angle of the transmitter it does not contain enough to locate an unambiguous azimuth angle in itself. Therefore by finding the correct Doppler for each azimuth, from the $Max Upper Envelope$ and the $| Min Lower Envelope |$, we can determine an unambiguous reference. This is done by evaluating the difference at each angle, and if the difference is negative the $| Min Lower Envelope |$ is used and if positive the $Max Upper Envelope$.

@@

From there the minimum method and the peak method can be compared to see if they agree on the 90\textdegree \space location. If this is not the case one will be chosen and the helicopter can fly in that direction. The resulting Doppler can then be measured, based on the physical movement of the helicopter or if the resulting envelope, to confirm the correct 90\textdegree \space location.

@@

\section{Transmitter Elevation Angle Estimation} \label{sec:transmitter_elevation_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
Using the information gathered from the azimuth estimation procedure, we can compute the elevation angle of the transmitter. In the process of gathering azimuth information, the maximum and minimum envelopes follow a sinusoidal pattern that contains a maximum Doppler information at approximately 180\textdegree \space and 0\textdegree \space azimuth angles. Using that maximum value, along with the correct Doppler and the assumption that the maximum will occur near the tip of the rotor blade, we can approximate the respected reflection radii off the blade as the azimuth angle changes. This is done through 
%reflection radius aproximation equation
\begin{equation}
	R_{Estimate} = BladeLength \left(\frac{f_{DopplerCorrect}(\theta_{Az})}{Max(f_{DopplerCorrect})}\right)
	\label{eqn:r_estimate}
\end{equation}

where $R_{Estimate}$ is the estimated reflection radius at a particular azimuth, $BladeLength$ is the length of a single blade or radius of the rotor, and $f_{DopplerCorrect}$ is the correct Doppler frequency for each azimuth angle $\theta_{Az}$.

@@

Now we find the maximum amount of Doppler that could be induced at that reflection radius location through
%max doppler at radius location
\begin{equation}
	f_{Doppler_{R Estimate Max}} = 2\frac{v_r f_c}{c}
	\label{eqn:max_doppler}
\end{equation}

where $f_{Doppler_{R Estimate Max}}$ is the maximum amount of Doppler that could happen at that location if the signal was reflected parallel to the direction of motion. $f_c$ is the center frequency of the signal, $c$ is the speed of light and $v_r$ is the radial velocity based of the $RPM$ defined by

%radial velocity
\begin{equation}
	v_r = \frac{2\pi}{60} RPM R_{Estimate}
	\label{eqn:v_radial}
\end{equation}

Then to find the elevation angle we look back at equation \ref{eq:theory_180_upper} swapping $MaxUpperEnvelope$ with $f_{DopplerCorrect(\theta_{Az})}$ because this is now over the azimuth range and swap $f_{DopplerMax}$ for $f_{Doppler_{R Estimate Max}}$ because we need the maximum doppler for the current reflection estimate. We then solve for the elevation angle

@@

where $\theta_{El}$ is the elevation angle of the transmitter. If the elevation angle is all the information needed equation \ref{eqn:elevation_angle} distills down to

@@

The error in elevation is caused by a limitation on the estimation. Where the Doppler shift is limited by the maximum amount Doppler that the blade rotation can produce shown in equation \ref{eqn:v_radial} by replacing $v_r$ with $v_rMax$ which is calculated by

%radial velocity
\begin{equation}
	v_{rMax} = \frac{2\pi}{60} RPM BladeLength
	\label{eqn:v_radial_max}
\end{equation}

then using that to find the maximum available Doppler through

@@

This limitation is shown by varying the elevation angle at the azimuth angle that provides the maximum Doppler shift, which is 180\textdegree. The maximum and minimum envelope data is shown in Figure \ref{fig:envelope_180deg}

%figure of 180 deg elevation sweep of envelope
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/results/Elevation_angle_envelopes_180deg_Azimuth.eps}
		\caption{Max and Min Envelope Frequencies vs Elevation Angle with Transmitter Azimuth Angle of 180\textdegree}
		\label{fig:envelope_180deg}
	\end{center}
\end{figure}

@@

Using the estimation method described above, Figure \ref{fig:angle_comparason_180deg} shows the elevation angle estimated versus the actual angle. The estimation method is accurate till elevation angles of 18\textdegree \space or 20\textdegree, where it becomes noisy. This is due to the inability of the envelope estimation to resolve the correct maximum Doppler for that range. By going back to Figure \ref{fig:tx_range_elevation_rel} we can see that as the transmitter moves farther and farther away the elevation angle is still decreasing but at a much smaller and smaller rate. So the difference between elevation angles closer to zero are more difficult to resolve.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/results/difference_pitch_noCorrection.eps}
		\caption{Uncorrected Envelope Difference vs. Transmitter Azimuth and Elevation angle of 45\textdegree with Blade pitch of 7.5\textdegree}
		\label{fig:diff_pitch_noCorrection}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/results/correct_dop_pitch_noCorrection.eps}
		\caption{Uncorrected Doppler vs. Transmitter Azimuth and Elevation angle of 45\textdegree with Blade pitch of 7.5\textdegree}
		\label{fig:dop_pitch_noCorrection}
	\end{center}
\end{figure}

Comparing the envelope difference of no pitch and pitched blade, in Figures \ref{fig:azimuth_estimation_difference} and \ref{fig:diff_pitch_noCorrection} respectively, we can see that the difference calculation of the pitch is a shifted up version of the one with no pitch. Therefore we can first correct this by re-centering the pitch difference according to

\begin{equation}
	A_{shift} = \frac{Max(Max Upper Envelope) - |Min(Min Lower Envelope)|}{2}
	\label{eqn:pitch_shift}
\end{equation}

where $A_{shift}$ is the amount that the difference calculation will be shifted down

\begin{equation}
	EnvelopeDifference = EnvelopeDifference - A_{shift}
	\label{eqn:diff_shift}
\end{equation}

Then with this knowledge $f_{DopplerCorrect}$ is recalculated and then is adjusted to the correct frequencies with

@@

Since $\theta_{Az}$ is unknown during the calculation of equation \ref{eqn:fd_pitch_correct} the correct starting phase can be determined by matching it to the $EnvelopeDifference$ and that the cosine should be one period in length. These corrections result in the percent azimuth error of 0\% and 7\% elevation angle error. The elevation angle percent error is within a useable range but could be decreased with further tuning to the preliminary pitch adjustment algorithm. When the pitch corrections are applied to a transmitter elevation angle of 26.56\textdegree \space the correction achieves better performance than the unpitched calculation with percent azimuth error of 0\% and percent elevation error of 0.114\%. This is still a preliminary correction method that requires more in depth analysis but from the initial results shows merit. The increased accuracy as the elevation angle decreases relates back to Figure \ref{fig:pitch_tx_elevation_angle} where as the elevation angle decreases the effect of pitch on the signal converges back to the unpitched results.

>>>>

\section{Introduction}
%introduce the results section and what processing will be done
The information gathered from the previous chapter is used to estimate the azimuth and elevation angles of the transmitter. First the azimuth angle will be estimated which will provide the information necessary for also calculating the elevation angle. The results of the estimation techniques are then analyzed to determine their accuracy.

@@

\section{Transmitter Azimuth Angle Estimation} \label{sec:transmitter_azimuth_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
The azimuth angle estimation technique is based on the data collected by rotating the helicopter platform in a clockwise manner, during which a Doppler profile and subsequent maximum and minimum values are calculated. After a complete rotation the Doppler maximum and minimum values produce a sinusoidal pattern as shown in Figure \ref{fig:tx_azimuth_angle_200} without knowing the starting azimuth angle. With this data we find an unambiguous angle that will set the coordinate space and provide an estimate of current azimuth angle. To do this the absolute value of the minimum Doppler profile data is superimposed onto the maximum Doppler profile data as shown in Figure \ref{fig:azimuth_estimation_max_vs_absMin}.

\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/results/Azimuth_angle_estimation_max_vs_absMin.eps}
		\caption{Max and Absolute value of Min Doppler Profile Frequencies vs Transmitter Azimuth Angle at an Elevation Angle of 45\textdegree}
		\label{fig:azimuth_estimation_max_vs_absMin}
	\end{center}
\end{figure}

From Figure \ref{fig:azimuth_estimation_max_vs_absMin} there are two points at which the maximums and the absolute value of the minimums cross. We then take the difference:

%formula for difference
\begin{equation}
	 Profile Difference = Max Doppler Profile - | Min Doppler Profile |
	 \label{eqn:difference}
\end{equation}

where $Max Doppler Profile$ is an array of the maximum value of each upper profile calculation, and $Min Doppler Profile$ is an array of the minimum value of each lower profile calculation. The result is an almost exact sinusoid centered around zero, shown in Figure \ref{fig:azimuth_estimation_difference}.

 \begin{figure}[b]
	\begin{center}
		\includegraphics[width=10cm]{images/results/Azimuth_angle_estimation_difference.eps}
		\caption{Difference between Max and Min Doppler Profile Calculations vs Transmitter Azimuth Angle at an Elevation Angle of 45\textdegree}
		\label{fig:azimuth_estimation_difference}
	\end{center}
\end{figure}

While Figure \ref{fig:azimuth_estimation_difference} provides information about the angle of the transmitter it does not contain enough to locate an unambiguous azimuth angle in itself. Therefore by finding the correct Doppler for each azimuth, from the $Max Doppler Profile$ and the $| Min Doppler Profile |$, we can determine an unambiguous reference. This is done by evaluating the difference at each angle, and if the difference is negative the $| Min Doppler Profile |$ is used and if positive the $Max Doppler Profile$.

@@

From there the minimum method and the peak method can be compared to see if they agree on the 90\textdegree \space location. If this is not the case one will be chosen and the helicopter can fly in that direction. The resulting Doppler can then be measured, based on the physical movement of the helicopter or if the resulting profile, to confirm the correct 90\textdegree \space location.

@@

\section{Transmitter Elevation Angle Estimation} \label{sec:transmitter_elevation_angle_estimation}
%formulate the estimation equation and discuss ambiguity considerations
Using the information gathered from the azimuth estimation procedure, we can compute the elevation angle of the transmitter. In the process of gathering azimuth information, the maximum and minimum profiles follow a sinusoidal pattern that contains a maximum Doppler information at approximately 180\textdegree \space and 0\textdegree \space azimuth angles. Using that maximum value, along with the correct Doppler and the assumption that the maximum will occur near the tip of the rotor blade, we can approximate the respected reflection radii by:
%reflection radius aproximation equation
\begin{equation}
	R_{Estimate} = r_b \left(\frac{f_{DopplerCorrect}(\theta_{Az})}{Max(f_{DopplerCorrect})}\right)
	\label{eqn:r_estimate}
\end{equation}

where $R_{Estimate}$ is the estimated reflection radius at a particular azimuth, $r_b$ is the length of a single blade or radius of the rotor, and $f_{DopplerCorrect}$ is the correct Doppler frequency for each azimuth angle $\theta_{Az}$.

@@

Now we find the maximum amount of Doppler that could be induced at that reflection radius location is:
%max doppler at radius location
\begin{equation}
	f_{Doppler_{R Estimate Max}} = 2\frac{v_r f_c}{c}
	\label{eqn:max_doppler}
\end{equation}

where $f_{Doppler_{R Estimate Max}}$ is the maximum amount of Doppler that could happen at that location if the signal was reflected parallel to the direction of motion. $f_c$ is the center frequency of the signal, $c$ is the speed of light and $v_r$ is the radial velocity based of the $\omega_r$ is:

%radial velocity
\begin{equation}
	v_r = \omega_r R_{Estimate}
	\label{eqn:v_radial}
\end{equation}

Then to find the elevation angle we look back at equation \ref{eq:theory_180_upper} swapping $MaxDopplerProfile$ with $f_{DopplerCorrect(\theta_{Az})}$ because this is now over the azimuth range and swap $f_{DopplerMax}$ for $f_{Doppler_{R Estimate Max}}$ because we need the maximum doppler for the current reflection estimate. We then solve for the elevation angle:

@@

where $\theta_{El}$ is the elevation angle of the transmitter. If the elevation angle is all the information needed equation \ref{eqn:elevation_angle} distills down to:

@@

The error in elevation is caused by a limitation on the estimation. Where the Doppler shift is limited by the maximum amount Doppler that the blade rotation can produce shown in equation \ref{eqn:v_radial} by replacing $v_r$ with $v_rMax$ which is:

%radial velocity
\begin{equation}
	v_{rMax} = \omega_r r_b
	\label{eqn:v_radial_max}
\end{equation}

then using $v_{rMax}$ to find the maximum available Doppler:

@@

This limitation is shown by varying the elevation angle at the azimuth angle that provides the maximum Doppler shift, which is 180\textdegree. The maximum and minimum profile data is shown in Figure \ref{fig:envelope_180deg}

%figure of 180 deg elevation sweep of envelope
\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/simulation/elevation_angle_max_doppler_180deg.eps}
		\caption{Max and Min Doppler Profile Frequencies vs Elevation Angle with Transmitter Azimuth Angle of 180\textdegree}
		\label{fig:envelope_180deg}
	\end{center}
\end{figure}

@@

Using the estimation method described above, Figure \ref{fig:angle_comparason_180deg} shows the elevation angle estimated versus the actual angle. The estimation method is accurate till elevation angles of 18\textdegree \space or 20\textdegree, where it becomes noisy. This is due to the inability of the profile estimation to resolve the correct maximum Doppler for that range. By going back to Figure \ref{fig:tx_range_elevation_rel} we can see that as the transmitter moves farther and farther away the elevation angle is still decreasing but at a much smaller and smaller rate. So the difference between elevation angles closer to zero are more difficult to resolve.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/results/difference_pitch_noCorrection.eps}
		\caption{Uncorrected Doppler Profile Difference vs. Transmitter Azimuth and Elevation angle of 45\textdegree with Blade pitch of 7.5\textdegree}
		\label{fig:diff_pitch_noCorrection}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=10cm]{images/results/correct_dop_pitch_noCorrection.eps}
		\caption{Uncorrected Pitch Doppler Profile vs. Transmitter Azimuth and Elevation angle of 45\textdegree with Blade pitch of 7.5\textdegree}
		\label{fig:dop_pitch_noCorrection}
	\end{center}
\end{figure}

Comparing the Doppler profile difference of no pitch and pitched blade, in Figures \ref{fig:azimuth_estimation_difference} and \ref{fig:diff_pitch_noCorrection} respectively, we can see that the difference calculation of the pitch is a shifted up version of the one with no pitch. Therefore we can first correct this by re-centering the pitch difference:

\begin{equation}
	A_{shift} = \frac{Max(Max Doppler Profile) - |Min(Min Doppler Profile)|}{2}
	\label{eqn:pitch_shift}
\end{equation}

where $A_{shift}$ is the amount that the difference calculation will be shifted:

\begin{equation}
	ProfileDifference = ProfileDifference - A_{shift}
	\label{eqn:diff_shift}
\end{equation}

Then $f_{DopplerCorrect}$ is recalculated and then adjusted to the correct frequencies:

@@

Since $\theta_{Az}$ is unknown during the calculation of equation \ref{eqn:fd_pitch_correct} the correct starting phase can be determined by matching it to the $ProfileDifference$ and that the cosine should be one period in length. These corrections result in the percent azimuth error of 0\% and 7\% elevation angle error. The elevation angle percent error is within a useable range but could be decreased with further tuning to the preliminary pitch adjustment algorithm. When the pitch corrections are applied to a transmitter elevation angle of 26.56\textdegree \space the correction achieves better performance than the unpitched calculation with percent azimuth error of 0\% and percent elevation error of 0.114\%. This is still a preliminary correction method that requires more in depth analysis but from the initial results shows merit. The increased accuracy as the elevation angle decreases relates back to Figure \ref{fig:pitch_tx_elevation_angle} where as the elevation angle decreases the effect of pitch on the signal converges back to the unpitched results.
